Narrative

### 2023-05-19 manuscript/introduction.qmd
improved first 2 paragraphs of introduction

Estimating heterogeneity of effect sizes is usually considered the second main purpose of meta-analysis, after estimating an average effect size. Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015). From a more statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes. However, true subpopulation effect sizes can refer to two different entities. Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015). Of these measurement artifacts, measurement error is present in any empirical effect size estimate and systematically attenuates effect sizes. It is usually measured as reliability (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF) and is relatively common to report in primary studies in psychology (Flake et al., 2017). However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015). This has consequences for heterogeneity estimates reported in many areas of psychology and in this article we use Monte-Carlo simulations to explore the consequences for heterogeneity estimates in meta-analysis from not correcting for unreliability.

Meta-analyses in psychology rarely correct for unreliability in primary study measurements. Wiernick and Dahlke (2020) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability. Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts. The exception is the subfield of Industrial-Organizational studies. Although at the meta-analysis level rather than study level, Cortina (2003) showed that about 80% of the 1,294 meta-analyses published across 59 meta-analytic studies in the Journal of Applied Psychology between 1978 - 1997 corrected for measurement artifacts, and Anguinis (2011) that around 50% of meta-analyses identified in studies published across several management journal between 1989 - 2009 corrected for unreliability of measurement.

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2, REF3), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.

>>>>

Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis in addition to estimating an average effect size. Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015). Heterogeneity is a primary outcome in meta-analysis because its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017), because it affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people, and because its presence changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity). As such, estimating heterogeneity without bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).

Meta-analytic estimates are highly dependent on the quality of the data in the primary studies. Unfortunately, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021). Such inattention to measurement can bias heterogeneity estimates. From a statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes. However, true subpopulation effect sizes can refer to two different entities. Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015). Of these measurement artifacts, measurement error (imperfect reliability of measurement) is present in any empirical effect size estimate and systematically attenuates effect sizes. However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015). This has consequences for heterogeneity estimates reported in many areas of psychology and in this article we use Monte-Carlo simulations to explore the consequences for heterogeneity estimates in meta-analysis from not correcting for unreliability.

Meta-analyses in psychology rarely correct for unreliability in primary study measurements. Wiernick and Dahlke (2020) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability. Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts. The exception is the subfield of Industrial-Organizational studies where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.
